Uncovering underlying high-level musical content in the time domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn deep audio embeddings with applications to MIR downstream tasks

### 2023-06-03 abstract/abstract.tex
Updates from Overleaf

Our cognitively-based approach for learning embeddings focuses on using full resolution data and preserving high-level musical information. Drawing upon our domain expertise, we developed robust transformations to encode heuristic musical concepts that should remain constant. This novel approach aims to reconcile music and machine learning, enhancing machine listening models' efficacy. Preliminary results suggest that our musically-informed technique has significant potential for boundary detection tasks and possibly nearly all MIR downstream tasks that are not purely sonic-based. Consequently, music-motivated embeddings appear promising and potentially adaptable to other tasks constrained by data scarcity.

\bigskip
Keywords: Music Information Retrieval; Music Structure Analysis; Deep Audio Embeddings, Aural Skills

>>>>

Our cognitively-based approach for learning embeddings focuses on using full-resolution data and preserving high-level musical information. Drawing upon our domain expertise, we developed robust transformations to encode heuristic musical concepts that should remain constant. This novel approach aims to reconcile music and machine learning, enhancing machine listening models' efficacy. Preliminary results suggest that our musically-informed technique has significant potential for boundary detection tasks and possibly nearly all MIR downstream tasks that are not purely sonic-based. Consequently, music-motivated embeddings appear promising and potentially adaptable to other tasks constrained by data scarcity.

\bigskip
Keywords: MIR; Music Structure Analysis; deep audio embeddings; audio representations; representation learning; embeddings; transfer learning; multi-task learning; multi-modal learning; aural Skills

### 2023-06-03 introduction/Subsections/Motivation.tex
Updates from Overleaf

Personal, academic, and industrial motivations drive this thesis.

This research is conducted in collaboration with \textit{Epidemic Sound AB (ES)} \cite{EpidemicSite}, and the \textit{Music Technology Group (MTG)} at Universitat Pompeu Fabra.

ES is a prominent Swedish company that provides a global library of over 40,000+ royalty-free soundtrack music and 90,000+ sound effects.

The MTG, based in Barcelona (Spain), specializes in developing new technologies related to sound and music, including music information retrieval, digital signal processing, computational models of music perception and cognition, and interactive music systems.

The collaboration with ES and the MTG aims to deepen our understanding of music and its fundamental structures, potentially leading to advancements and benefits for the company's technical products, services, and MIR academic research.

\subsubsection{Personal motivation}

As a music enthusiast, musician, and scholar, I am passionately committed to exploring the foundational elements of musical composition, investigating the diverse techniques employed in creating musical works, and deciphering the intricate relationships among them.

This passion fuels my quest for knowledge on extracting valuable information embedded within music in any domain, being my main interest in the audio and symbolic domains \cite{Crestel2018AOrchestration}\cite{WeilAutomaticSignals}, contributing to developing more sophisticated and effective music-related technologies and solutions in the industrial landscape.

\subsubsection{Industrial motivation}

While it is true that the theoretical foundations of machine learning can be traced back to the pioneering work of figures such as Alan Turing and Claude Shannon in the mid-twentieth century, we are currently experiencing a period of significant breakthroughs in AI research \cite{Vaswani2017AttentionNeed}. These breakthroughs have facilitated the implementation of AI in widely used products \cite{OpenAI2023GPT-4Report}. Many researchers and corporations are keen to stay abreast of these developments and contribute to this exciting and rapidly evolving field.

The emergence of self-supervised models that learn embedding spaces for retrieving musical content from audio signals has unveiled new opportunities. By capturing high-level semantic information from raw audio data in an unsupervised manner, these models can potentially revolutionize various aspects of the music industry.

In the industrial and corporate realm, the embedding spaces generated by these models can be utilized to create all sorts of valuable products and user experiences for content creators. These advancements ultimately enhance our interaction with music in the digital age, promoting innovation and collaboration across the music industry.

>>>>

Driven by personal, academic, and industrial motivations, this thesis is a product of a collaboration between \href{https://www.epidemicsound.com/}{Epidemic Sound} (ES) and the \href{https://www.upf.edu/web/mtg}{Music Technology Group} (MTG) at Universitat Pompeu Fabra. ES, a leading Swedish company, holds and curates an expansive global library of 40,000+ royalty-free music tracks and 90,000+ sound effects. MTG, based in Barcelona, Spain, excels in developing innovative technologies related to sound and music, such as music information retrieval, digital signal processing, interactive music systems, and computational musicology.

This partnership seeks to enhance our understanding of music's fundamental structures, potentially improving ES's technical products, services, and advancements in Music Information Retrieval (MIR) research.

As a music enthusiast, musician, and scholar, I am deeply committed to exploring the foundational elements of musical composition, investigating various creation techniques, and deciphering their intricate relationships. This passion drives my endeavor to extract valuable information embedded within music across any domain, particularly in audio and symbolic domains.

We are witnessing significant breakthroughs in AI research \cite{Vaswani2017AttentionNeed}, building on the theoretical foundations of pioneers such as Alan Turing and Claude Shannon. These advancements have resulted in the widespread application of AI in popular products \cite{OpenAI2023GPT-4Report}, prompting researchers and corporations to stay informed and contribute to this rapidly evolving field. The advent of self-supervised models that learn embedding spaces for retrieving musical content from audio signals presents fresh opportunities. By autonomously extracting high-level semantic information from raw audio data, these models could transform various facets of the music industry. In the industrial realm, these embedding spaces can be harnessed to devise valuable products and enrich user experiences for content creators, thereby fostering innovation and collaboration across the music industry.

### 2023-06-03 introduction/Subsections/Objectives.tex
Updates from Overleaf

The study aims to develop a unified embedding, $e(M)$, for a music piece $M$ that integrates high-level elements like melody contour, harmony relationships, and other comprehensive components, arranging the embedding space to emphasize clustering of these sophisticated musical content without any consideration for sonic quality attributes. Utilizing a blend of Music Information Retrieval (MIR) computational methodologies and musicological perspectives, the research will explore musical compositions from various styles and genres.

>>>>

The study aims to develop a unified numerical understanding (or embedding \footnote{A learned representation or embedding is a numerical output -usually a fixed-size vector- produced by a machine-learning model from a low-level input. Good representations should be versatile across various tasks and require limited supervision \cite{Turian2022HEAR:Representations}.}) for a music piece that integrates high-level elements like melody contour, harmony relationships, and other comprehensive components such as the relationships between elements unfolding through time, arranging the embedding space to emphasize the clustering of these sophisticated musical content without considering sonic qualities or attributes.

The research will utilize MIR computational methodologies and musicological perspectives to explore musical compositions from various styles and genres.

### 2023-06-03 introduction/Subsections/Previous work.tex
Updates from Overleaf

Neural networks have recently facilitated the development of low-dimensional embeddings that capture essential musical features, enabling the computation of various task-specific aspects.

These aspects include the utilization of multi-level music segmentation through deep audio embeddings to replace manually engineered features \cite{SalamonDeepSegmentation}, the enhancement of environmental sound classification \cite{Kim2020OneStrategies}, the improvement of vocal-centric music tasks using cross-domain audio embedding \cite{Kim2021LearningLoss}, the enhancement of audio classification through a combination of task-specific and pre-trained features \cite{Hung2022Feature-informedClassification}, the creation of a music similarity search engine specifically for video producers \cite{epidemic}, the improved performance in Music Emotion Recognition (MER) tasks \cite{KohComparisonRecognition}, embedding songs for similarity comparison based on artist identification \cite{contentmusicsimtriplet2020}, the resolution of the cross-modal text-to-music retrieval problem \cite{WonEmotionStories}, and automated music rearrangement \cite{Stoller2018IntuitiveTransitions}\cite{Plachouras2023MusicSegmentation}.

On top of that and returning to musicological influences, the middle ground in Schenkerian analysis can be compared to the mesostructure in music. Analogous to the mesostructure, it represents the intermediate-level musical patterns that bridge the micro and macro structures. This element is often overlooked when applying deep learning to music. As emphasized by \cite{Mesostructures2023}, models that address this mesostructure level could significantly enrich tasks such as music analysis, composition, and retrieval, providing a more holistic understanding of music, similar to the insights gained from analyzing the middle-ground in Schenkerian analysis. \cite{Introduction_to_Schenkerian_Analysis}

Moreover, deep audio embeddings offer the advantage of transferability. Once trained, these latent representations can be utilized as a starting point for various other tasks, thus saving computational resources and time compared to training a model from scratch \cite{HamelTransferSimilarity}.

Given their demonstrated success in the literature and their potential for transfer learning, exploring deep audio embeddings for MIR downstream tasks seems worthwhile.

>>>>

Recent advancements in neural networks have enabled the development of low-dimensional embeddings which encapsulate critical musical attributes, thus facilitating the computation of diverse task-specific elements.

Such tasks encompass developing an audio embedding method that excels across a wide array of applications without fine-tuning \cite{Turian2022HEAR:Representations}, deploying deep audio embeddings for multi-level music segmentation \cite{SalamonDeepSegmentation}, improving environmental sound classification \cite{Kim2020OneStrategies}, enhancing vocal-centric music tasks via cross-domain audio embedding \cite{Kim2021LearningLoss}, combining task-specific and pre-trained features for improved audio classification \cite{Hung2022Feature-informedClassification}, creating a music similarity search engine tailored for video producers \cite{epidemic}, enhancing Music Emotion Recognition (MER) performance \cite{KohComparisonRecognition}, investigating the efficacy of speaker recognition pre-trained model embeddings \cite{lightweight}, embedding songs for similarity comparison via artist identification \cite{contentmusicsimtriplet2020}, solving the cross-modal text-to-music retrieval issue \cite{WonEmotionStories}, and facilitating automated music rearrangement \cite{Stoller2018IntuitiveTransitions, Plachouras2023MusicSegmentation}.

Furthermore, deep learning applications in music often overlook the intermediate-level musical patterns, analogous to the 'mesostructure' in music and the middle ground in Schenkerian analysis. As highlighted by \cite{Mesostructures2023}, addressing this mesostructure could significantly improve tasks like music analysis, composition, and retrieval by providing a more holistic understanding of music, comparable to insights from Schenkerian analysis \cite{Introduction_to_Schenkerian_Analysis}.

Additionally, deep audio embeddings offer the benefit of transferability, acting as a starting point for various tasks, thus saving computational resources and time as opposed to training a model from scratch \cite{HamelTransferSimilarity}.

However, it is still an open question whether a general-purpose audio representation can effectively mimic human hearing \cite{Turian2022HEAR:Representations}.

Considering their proven efficacy in existing literature and their potential for transfer learning, it appears promising to further investigate deep audio embeddings for Music Information Retrieval (MIR) downstream tasks.

### 2023-06-03 methods/subsections/implementation.tex
Updates from Overleaf

Our approach could be broadly categorized as cognitive modeling, as it strives to mimic human cognitive processes and problem-solving through a computerized model. We champion exposure-based learning in music and active engagement with an array of musical styles, genres, techniques, and learning methods to foster well-rounded musical proficiency and efficient performance with novel data in practical applications.

@@

Our experiments can benefit from incorporating time-domain processing, which allows for more natural handling of temporal patterns and sequences in the data while avoiding windowing artifacts. While mel-spectrogram-based approaches have proven useful in various audio-related machine-learning tasks, their limitations lie in their representation of perceptual similarity. Spectrograms capture the distribution of frequencies over time, and small differences in spectrograms generally indicate high perceptual similarity from a human listener's perspective. However, there are cases where two audio signals can have significantly different spectrograms but still be perceived as musically similar. This discrepancy is due to the complexity of human auditory perception, which encompasses temporal patterns, phase relationships between frequencies, and higher-level musical structures. Therefore, relying solely on spectrogram distance to measure high-level music similarity may not consistently align with human perceptual judgments \cite{Kim2020OneStrategies}\cite{Mesostructures2023}.

@@

We posit that high-level musical content unfolds over time; therefore, we argue that the temporal structure of the negative images related to our anchor should be disrupted. While maintaining similar sonic qualities, the content should be rendered unintelligible.

\begin{enumerate}
\item Calculate the minimum and maximum audio chunk lengths in samples:
\begin{equation}
l_{min} = t_{min} \times S
\end{equation}
\begin{equation}
l_{max} = t_{max} \times S
\end{equation}
\item Generate random audio chunk lengths $l_1, l_2, \ldots, l_{n-1}$ from the uniform distribution on the interval $[l_{min}, l_{max}]$. Calculate the final audio chunk length as:
\begin{equation}
l_n = L_A - \sum_{i=1}^{n-1} l_i
\end{equation}
where $L_A$ is the length of the anchor signal in samples.
\item Split the anchor signal $A$ into audio chunks $C_1, C_2, \ldots, C_n$ according to the calculated audio chunk lengths.
\item Shuffle the audio chunks randomly to get the permuted slices $C_{\sigma(1)}, C_{\sigma(2)}, \ldots, C_{\sigma(n)}$, where $\sigma$ is a random permutation of indices from $1$ to $n$.
\item Concatenate the shuffled audio chunks to generate the negative signal:
\begin{equation}\label{eq:negative_signal}
N = C_{\sigma(1)} \oplus C_{\sigma(2)} \oplus \ldots \oplus C_{\sigma(n)}
\end{equation}
\end{enumerate}

@@

As previously stated, the goal of minimizing this loss function is to learn discriminative embeddings, where similar examples are grouped closely together while dissimilar examples are placed farther apart in the embedding space.

>>>>

Our approach could be broadly categorized as cognitive modeling, as it strives to mimic human cognitive processes and problem-solving through a computerized model. We champion exposure-based learning in music and active engagement with musical styles, genres, techniques, and learning methods to foster well-rounded musical proficiency and efficient performance with novel data in practical applications.

@@

Conventional transformations, such as log-scaled Mel spectrograms or Mel Frequency Cepstral Coefficients, capture key features but may obscure some information inherent in raw audio data.

Raw audio data offers unique advantages. Primarily, it ensures the retention of the original signal, which could reveal novel findings, and facilitates the direct extraction of features through advanced deep learning models \cite{learning}. However, it is not without challenges: high sample rates result in high dimensionality, demanding substantial computational resources. Additionally, raw audio is noise-sensitive, complicating the extraction process. Features derived from raw audio also often lack the interpretability of traditional ones \cite{Schindler2020DeepTutorial}. Hence, the choice of input data depends on the specific task, available computational resources, and the balance between data preservation and computational efficiency.

Time-domain processing, a feature of our experiments, naturally handles temporal patterns and sequences in data, thereby avoiding windowing artifacts. Although Mel-spectrogram-based methods are effective for various audio-related machine learning tasks, their limitations lie in representing perceptual similarity. Spectrograms capture frequency distribution over time. Yet, the complexity of human auditory perception—encompassing temporal patterns, phase relationships between frequencies, and higher-level musical structures—means that musically similar sounds can have distinct spectrograms. This discrepancy implies that using spectrogram distance alone for measuring high-level music similarity may not always align with human perceptions \cite{Kim2020OneStrategies}\cite{Mesostructures2023}.

@@

Circling back to our premises and assumptions, we posit that high-level musical content unfolds over time; therefore, we argue that the temporal structure of the negative images, in contrast to our anchor, should be disrupted. While maintaining similar sonic qualities, the content should be rendered unintelligible.

\begin{enumerate}
\item Calculate the minimum and maximum audio chunk lengths in samples:
\begin{equation}
l_{min} = t_{min} \times S
\end{equation}
\begin{equation}
l_{max} = t_{max} \times S
\end{equation}

The minimum duration $t_{min}$ is set to 0.05 seconds, and the maximum duration $t_{max}$ is set to 1 second. This range is chosen thoughtfully to strike a balance between the two factors. On the one hand, it is above the just noticeable difference (JND) threshold, the smallest change in a stimulus that can be perceived. On the other hand, it is short enough to maintain a reasonably-sized window to avoid discernible musical content \cite{Fastl2007Just-NoticeableChanges}.

\item Generate random audio chunk lengths $l_1, l_2, \ldots, l_{n-1}$ from the uniform distribution on the interval $[l_{min}, l_{max}]$. Calculate the final audio chunk length as:
\begin{equation}
l_n = L_A - \sum_{i=1}^{n-1} l_i
\end{equation}
where $L_A$ is the length of the anchor signal in samples.
\item Split the anchor signal $A$ into audio chunks $C_1, C_2, \ldots, C_n$ according to the calculated audio chunk lengths.
\item Shuffle the audio chunks randomly to get the permuted slices $C_{\sigma(1)}, C_{\sigma(2)}, \ldots, C_{\sigma(n)}$, where $\sigma$ is a random permutation of indices from $1$ to $n$.
\item Concatenate the shuffled audio chunks to generate the negative signal:
\begin{equation}\label{eq:negative_signal}
N = C_{\sigma(1)} \oplus C_{\sigma(2)} \oplus \ldots \oplus C_{\sigma(n)}
\end{equation}
\end{enumerate}

@@

As previously stated, the goal of minimizing this loss function is to learn discriminative embeddings, where similar examples are grouped closely together. In contrast, dissimilar examples are placed farther apart in the embedding space.

While some packages can be found in the MIR literature \cite{auraloss}, we wrote our own PyTorch \cite{Paszke2019PyTorch:Library} implementation for the sake of our experiments.

### 2023-06-03 results/subsections/tables.tex
Updates from Overleaf

@@

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{4.5cm}XXXXX}
\toprule
\thead{\centering\textbf{Year, Authors, Ref}} & \thead{\centering\textbf{Algorithm}} & \thead{\centering\textbf{Input}} & \thead{\centering\textbf{Method}} & \thead{\centering\textbf{F-measure \footnote{The F-measure is a widely used boundary detection metric computed by comparing the predicted and ground truth boundaries. This metric, ranging from 0 to 1, represents the harmonic mean of precision and recall, providing a balanced view of under-segmentation and over-segmentation. Precision calculates the percentage of correctly detected boundaries, whereas recall assesses the percentage of actual boundaries detected correctly. Given the potential inaccuracies in human annotations and prediction errors, the F-measure tolerates minor discrepancies between predicted and ground truth boundaries. This tolerance level can be adjusted, enabling the metric to consider a predicted boundary as correct if it's within the set tolerance window of a ground truth boundary \cite{Turnbull2007ABOOSTING}.}}} \\
\midrule
\addlinespace
2012, Kaiser et al., [27]& KSP2 & SSM & Novelty measure  & 0.286 \\
\addlinespace
2013, McFee \& Ellis, [20] & MP2 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
2014, Nieto \& Bello, [28] & NB1 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
2015, Cannam et al., [29] & CC1 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
2016, Nieto, [30] & ON2 & Constant-Q Transform Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
2017, Cannam et al., [29] & CC1 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
2007, Turnbull et al., [19] & - & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
2011, Sargent et al., [34] & - & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
2014, Ullrich et. al, [22] & - & MLS & CNN  & 0.465 \\
\addlinespace
2015, Grill \& Schlüter, [4] & - & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
2015, Grill \& Schlüter, [5] & - & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
2017, Hadria \& Peeters, [35] & - & MLS, SSLMs & CNN  & 0.291 \\
\bottomrule
\end{tabularx}
\caption{Your Caption}
\label{tab:my_label}
\end{table}

>>>>

METRICS

\begin{itemize}
    \item \textbf{Precision}: Precision measures the proportion of correctly detected boundaries out of the total estimated boundaries. It quantifies the accuracy of the boundary detection algorithm. It is calculated as:

    \[
    \text{Precision} = \frac{\text{True Positives}}{\text{True Positives} + \text{False Positives}}
    \]

    High precision indicates a low rate of false positives, meaning the algorithm correctly identifies boundaries without many incorrect detections.

    \item \textbf{Recall}: Recall measures the proportion of correctly detected boundaries out of the total reference boundaries. It quantifies the completeness or sensitivity of the boundary detection algorithm. It is calculated as:

    \[
    \text{Recall} = \frac{\text{True Positives}}{\text{True Positives} + \text{False Negatives}}
    \]

    High recall indicates a low rate of false negatives, meaning that the algorithm successfully detects most of the true boundaries without missing many.

    \item \textbf{F-measure}: The F-measure is a widely used boundary detection metric computed by comparing the predicted and ground truth boundaries. This metric, ranging from 0 to 1, represents the harmonic mean of precision and recall, providing a balanced view of under-segmentation and over-segmentation. Precision calculates the percentage of correctly detected boundaries, whereas recall assesses the percentage of actual boundaries detected correctly \cite{Turnbull2007ABOOSTING}. The F-measure is calculated as follows:

    \[
    \text{F-measure} = \frac{2 \times \text{Precision} \times \text{Recall}}{\text{Precision} + \text{Recall}}
    \]
\end{itemize}

@@

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{4.5cm}XXXXX}
\toprule
\thead{\centering\textbf{Year, Authors, Ref}} & \thead{\centering\textbf{Algorithm}} & \thead{\centering\textbf{Input}} & \thead{\centering\textbf{Method}} & \thead{\centering\textbf{F-measure}} \\
\midrule
\addlinespace
2012, Kaiser et al., [27]& KSP2 & SSM & Novelty measure  & 0.286 \\
\addlinespace
2013, McFee \& Ellis, [20] & MP2 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
2014, Nieto \& Bello, [28] & NB1 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
2015, Cannam et al., [29] & CC1 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
2016, Nieto, [30] & ON2 & Constant-Q Transform Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
2017, Cannam et al., [29] & CC1 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
2007, Turnbull et al., [19] & - & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
2011, Sargent et al., [34] & - & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
2014, Ullrich et. al, [22] & - & MLS & CNN  & 0.465 \\
\addlinespace
2015, Grill \& Schlüter, [4] & - & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
2015, Grill \& Schlüter, [5] & - & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
2017, Hadria \& Peeters, [35] & - & MLS, SSLMs & CNN  & 0.291 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State of the art table.]{Your Caption. Given the potential inaccuracies in human annotations and prediction errors, the F-measure tolerates minor discrepancies between predicted and ground truth boundaries. This tolerance level can be adjusted, enabling the metric to consider a predicted boundary as correct if it's within the set tolerance window of a ground truth boundary.}
\label{tab:my_label}
\end{table}
